---
title: "Statictical Modeling and Advanced Regression Analyses"
subtitle: "R Tutorials"
date: now
theme:
  - cosmo
  - quartosupport2.scss
highlight-style: monochrome
linkcolor: gray
notebook-links: false
date-format: long
bibliography: references.bib
csl: quartosupport3.csl
author:
  - name: Holger Sennhenn-Reulen\textsuperscript{\orcidlink{0000-0002-4782-4387}} \linebreak \textsuperscript{Northwest German Forest Research Institute (NW-FVA), Germany.}
    orcid: 0000-0002-4782-4387
    corresponding: true
    email: holger dot sennhenn at nw minus fva dot de
    affiliation: "Northwest German Forest Research Institute (NW-FVA)"
    roles:
      - Statistician
      - Researcher
plain-language-summary: |
  This document shows R code and results. 
key-points:
  - kp1
  - kp2
---

<!-- ```{r} -->
<!-- if(knitr::is_latex_output()) { -->
<!--   knitr::opts_chunk$set(dev = 'tikz') -->
<!--   options(tikzDocumentDeclaration = "\\documentclass{article}\n\\usepackage[lining]{ebgaramond}") -->
<!-- } -->
<!-- ``` -->

`r if (knitr::is_latex_output()) '\\clearpage'`

# Software

We use the statistical software environment *R* [@RCoreTeam2024], and R add-on packages *ggplot2* [@Wickham2016], and *arm* [@GelmanSu2024]. 

<!-- and *plyr* [@Wickham2011].  -->
<!-- *colorspace* [@StaufferEtAl2009] -->
<!-- **cowplot** [@Wilke2023] -->

This document is produced using *Quarto* [@AllaireEtAl2024]. 
<!-- Graphics in the PDF version are rendered using R add-on package *tikzDevice* [@SharpsteenBracken2023].  -->

## Organize R Session

```{r}
#| eval: true
#| echo: true
rm(list = ls())
library("ggplot2")
```

# Linear Regression Model

Data are simulated according to the equations given in the lecture slides[^footnoteA]:

[^footnoteA]: For two covariates `x_1` and `x_2`.

```{r}
#| eval: true
#| echo: true
set.seed(123)
N <- 500
df <- data.frame(x_1 = runif(n = N), 
                 x_2 = runif(n = N))
(beta_0 <- rnorm(n = 1, mean = 1, sd = .1))
(beta_x_1 <- rnorm(n = 1, mean = 1, sd = .1))
(beta_x_2 <- rnorm(n = 1, mean = -.5, sd = .1))
(sigma <- rgamma(n = 1, shape = 1, rate = 4))
df$mu <- beta_0 + beta_x_1 * df$x_1 + beta_x_2 * df$x_2
df$y <- df$mu + rnorm(n = N, mean = 0, sd = sigma)
```

```{r}
#| eval: true
#| echo: true
#| label: fig-plotA
#| fig-cap: "Scatterplot of the two simulated covariates `x_1` and `x_2` - each from the uniform distribution between $0$ and $1$."
ggplot(data = df, aes(x = x_1, y = x_2)) + 
  geom_point()
```

`r if (knitr::is_latex_output()) '\\clearpage'`

# References {-}

::: {#refs}
:::