---
date: "Oktober 26, 2021 (Version 0.1)"
#author: "Holger Sennhenn-Reulen"
title: "**Introduction to R: Session 03**"
author: "Holger Sennhenn-Reulen^[Private webpage: [uncertaintree.github.io](https://uncertaintree.github.io)], Nordwestdeutsche Forstliche Versuchsanstalt (NW-FVA)"
publisher: "Me"
site: bookdown::bookdown_site
documentclass: scrartcl
fontsize: 10pt
bibliography: references.bib
# biblio-style: "apalike"
link-citations: yes
output:
  bookdown::gitbook:
    highlight: monochrome
    css: style_gitbook.css
    citation_package: default
    anchor_sections: FALSE
    split_by: "none"
    config:
      toc:
        collapse: subsubsection
        scroll_highlight: yes
        before: null
        after: null
      toolbar:
        position: fixed
      edit : null
      download: [["intro_to_R_session_01.pdf", "PDF"]]
      search: no
      fontsettings: no
      info: no
      sharing: no
  bookdown::pdf_book:
    pandoc_args: --top-level-division=section
    citation_package: natbib
    highlight: monochrome
    includes:
      in_header: "preamble_gitbook.tex"
---

<!--`r if (knitr::is_html_output()) '# About this course{-}'`
`r if (knitr::is_html_output()) 'This course introduces basic R-programming skills needed for applied statistical modeling in R.'`-->

`r if (knitr::is_html_output()) 'All contents are licensed under [CC BY-NC-ND 4.0](https://creativecommons.org/licenses/by-nc-nd/4.0/).'`
`r if (knitr::is_latex_output()) '\\vfill'`
`r if (knitr::is_latex_output()) '\\begin{center}'`
`r if (knitr::is_latex_output()) '\\begin{minipage}{.6\\textwidth}'`
`r if (knitr::is_latex_output()) '\\textsf{All contents are licensed under CC BY-NC-ND 4.0 \\href{https://creativecommons.org/licenses/by-nc-nd/4.0/}{(Link)}.}'`
`r if (knitr::is_latex_output()) '\\end{minipage}'`
`r if (knitr::is_latex_output()) '\\end{center}'`
`r if (knitr::is_latex_output()) '\\vfill'`
`r if (knitr::is_latex_output()) '\\clearpage'`

# Data {-}

## Drought {-}

Data basis: Fischer, R., Dobbertin, M., Granke, O., et al., 2006. The condition of forests in Europe. 2006
Executive report. UNECE, Hamburg.

```{r}
bair <- c(.505, .648, .523, .426, .64, .5, .257, .866, .434, .368, .54, .923, .702, 
          .615, 1.013, .807, .262, .887, 1.281, 1.125, .99, 1.2, .983, .697, .606, 
          .718, .48, .822, .944, .77, 1.036, 1.23, .68, .985)
elev <- c(335, 460, 480, 515, 540, 650, 680, 715, 730, 835, 860, 960,
          1020, 1025, 1100, 1150, 1150, 1170, 1190, 1350, 1400, 1500, 1540,
          475, 480, 507.5, 580, 750, 780, 800, 1025, 1100, 1150, 1200)
species <- c("Spruce", "Spruce", "Spruce", "Spruce", "Spruce", "Spruce",
             "Spruce", "Spruce", "Spruce", "Spruce", "Spruce", "Spruce", "Spruce",
             "Spruce", "Spruce", "Spruce", "Spruce", "Spruce", "Spruce", "Spruce",
             "Spruce", "Spruce", "Spruce", "Beech", "Beech", "Beech", "Beech",
             "Beech", "Beech", "Beech", "Beech", "Beech", "Beech", "Beech")
drought <- data.frame(bair = bair,
                      elev = elev,
                      species = species)
summary(drought)
```
For further context information, another source working on and interpereting this data (p. 202-203): 

Matthias Dobbertin, Markus Neumann, Hans-Werner Schroeck, Chapter 10 - Tree Growth Measurements in Long-Term Forest Monitoring in Europe,
Editor(s): Marco Ferretti, Richard Fischer,
Developments in Environmental Science,
Elsevier,
Volume 12,
2013,
Pages 183-204,
[https://doi.org/10.1016/B978-0-08-098222-9.00010-8](https://doi.org/10.1016/B978-0-08-098222-9.00010-8)

### Frost {-}

<!--Datenbasis: Deutscher Wetterdienst, hier dargestellte Werte selbst aus Tageseinzelwerten erzeugt.-->
Data basis: Deutscher Wetterdienst, values shown here were generated over individual values by myself.

Direct download links for data basis (Stations Id 1691, Goettingen): 

- [historical data)](ftp://opendata.dwd.de/climate_environment/CDC/observations_germany/climate/daily/kl/historical/tageswerte_KL_01691_19270101_20201231_hist.zip)
- [recent data)](ftp://opendata.dwd.de/climate_environment/CDC/observations_germany/climate/daily/kl/recent/tageswerte_KL_01691_akt.zip)

Some definitions:

- Budburst is estimated based on first day with degree days > 220 (begin counting on March, 20). 
- End of 1st development stage is estimated based on first day with degree days > 320 (begin counting on March, 20). 
- Definition frost event: $\min\left(\text{Temp}_{50\text{cm}}\right) < -1.95°\text{C}$

```{r}
frost <- data.frame(year = 1947:2021, 
                    n_frost = c(0, 0, 2, 0, 0, 0, 0, 0, 0, 0, 2, 0, 0, 0, 0, 1, 
                                0, 0, 2, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 
                                0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 
                                2, 0, 0, 0, 0, 0, 0, 0, 0, 1, 2, 0, 0, 0, 0, 0, 
                                3, 2, 0, 0, 0, 0, 0, 0, 1, 5, 0), 
                    bud_burst = as.Date(c(-19230, -18867, -18503, -18127, -17758, -17408, -17034, 
                                          -16661, -16293, -15929, -15566, -15202, -14847, -14475, 
                                          -14121, -13744, -13384, -13017, -12647, -12291, -11917, 
                                          -11563, -11191, -10821, -10462, -10092, -9720, -9361, 
                                          -8997, -8635, -8261, -7896, -7530, -7164, -6808, -6436, 
                                          -6078, -5705, -5347, -4981, -4619, -4254, -3883, -3524, 
                                          -3145, -2788, -2437, -2060, -1694, -1322, -958, -602, 
                                          -237, 124, 499, 864, 1222, 1592, 1957, 2321, 2681, 3055, 
                                          3408, 3789, 4137, 4513, 4877, 5234, 5610, 5976, 6345, 
                                          6691, 7074, 7435, 7812), 
                                        origin = as.Date("2000-01-01")),
                    end_1st_dev_stage = as.Date(c(-19222, -18859, -18489, -18118, -17746, -17397, 
                                                  -17026, -16650, -16280, -15921, -15552, -15192, 
                                                  -14837, -14464, -14104, -13726, -13370, -13006, 
                                                  -12633, -12281, -11905, -11545, -11180, -10808, 
                                                  -10455, -10078, -9710, -9349, -8984, -8623, -8248, 
                                                  -7886, -7521, -7151, -6799, -6427, -6068, -5691, 
                                                  -5338, -4972, -4601, -4246, -3875, -3513, -3131, 
                                                  -2780, -2426, -2050, -1679, -1311, -944, -594, 
                                                  -225, 132, 510, 873, 1235, 1608, 1972, 2332, 2694, 
                                                  3067, 3422, 3802, 4152, 4525, 4891, 5250, 5623, 
                                                  5988, 6354, 6703, 7086, 7450, 7824), 
                                                origin = as.Date("2000-01-01")))
frost$may1st <- as.Date(paste0(frost$year, "-05-01"))
frost$bud_burst_days_since_may1st <- julian(frost$bud_burst, origin = as.Date("2000-01-01")) - 
  julian(frost$may1st, origin = as.Date("2000-01-01"))
frost$end_1st_dev_stage_days_since_may1st <- julian(frost$end_1st_dev_stage, 
                                                    origin = as.Date("2000-01-01")) - 
  julian(frost$may1st, origin = as.Date("2000-01-01"))
summary(frost)
```

`r if (knitr::is_latex_output()) '\\clearpage'`

<!--
# Beispieldaten {-}

```{r, echo = F, warning = F, message=FALSE}
library("lmfor")
library("plyr")
data(spati2)
d <- spati2
rm(spati2)
d_breaks_cut <- quantile(d$d, probs = seq(0, 1, by = 0.05))
d$d_cut <- cut(d$d, breaks = d_breaks_cut, include.lowest = T)
dd <- ddply(d, c("d_cut"), summarise, h_mean = mean(h),
            h_q25 = quantile(h, probs = 0.25),
            h_q75 = quantile(h, probs = 0.75))
dd$d_lb <- d_breaks_cut[-length(d_breaks_cut)]
dd$d_ub <- d_breaks_cut[-1]
dd$b_mean <- apply(dd[, c("d_lb", "d_ub")], MAR = 1, FUN = mean)
copy_original_par_mar <- par()$mar
par(mar = c(5.1, 4.1, 0.1, 2.1))
```
-->

# Objectives of control structures.

**'Automation' of the repetition of structurally identical commands.**

- Repetition of a command -- with parameter / quantities remaining the same or changing -- with a predetermined or flexible number of repetitions.
- Conditional execution of various tasks.
- Generalization of tasks by defining functions.
- Combination of information in objects.

`r if (knitr::is_latex_output()) '\\clearpage'`

# Logical comparisons.

Command   `TRUE` if:
--------- --------------
`==`      Equality
`! =`     Inequality
`>`, `>=` left side greater than (or equal to) the right side
`<`, `<=` left side less than (or equal to) the right side
`% in%`   Is left side in vector on right side?

- `all ()` returns `TRUE` if all elements of the vector are `TRUE`.
- `any ()` returns `TRUE` if at least one element of the vector is `TRUE`.
- `is.na()` and `is.null()` return `TRUE` if the respective object (e.g. element of a vector) is `NA` or `NULL`.
- a logical value can be negated with a preceding `!` (e.g. `!TRUE` is` FALSE`)
- `which()` returns the index set (as an integer vector) if the logical comparison resulted in `TRUE`.

## Exercises

```{r, results='hide'}
is.na(drought$bair)
any(is.na(drought$bair))
drought$bair > 0
all(drought$bair > 0)
drought$bair > 1
any(drought$bair > 1)
all(drought$bair > 1)
which(drought$bair > 1)
drought$bair[which(drought$bair > 1)]
(tmp <- round(drought$bair, 1))
c(.8, 1.2) %in% tmp
c(.8, 1.2) %in% drought$bair
which(tmp %in% c(.8, 1.2))
drought$bair[which(tmp %in% c(.8, 1.2))]
tmp <- c(drought$bair[1:5], NA)
all(tmp > 0)
any(is.na(tmp))
which(is.na(tmp))
all(tmp[-which(is.na(tmp))] > 0)
mean(tmp)
mean(tmp, na.rm = T)
```

`r if (knitr::is_latex_output()) '\\clearpage'`

# Conditional execution

## 'if-else'

**Usage:**
```{r, eval=F}
if (Condition) {
  ... ## Commands if Condition is TRUE
} else {
  ... ## Commands if Condition is FALSE
}
```

- `TRUE` or `FALSE` condition necessary.
- 'if-else'-sequences can be nested within one another.

**Example together with the next topic.**

## 'for'-loops

'for' loops often offer a simple and pragmatic way to complete steps in data management / preparation.

**Usage:**
```{r, eval=F}
for (index in vector) {
  ... index ... ## Command that in some form depend on index.
}
```

- New object `index` runs all elements in `vector`.
- `index` remains constant during `... index ...`
- `index` jumps to the next (if available) value of `vector` after running through `... index ...`.
- `index` takes each value of `vector` once.
- The number of iterations of `... index ...` is determined by the length of `vector`.

## Example of a `for` loop

```{r}
tmp1 <- frost$bud_burst_days_since_may1st
tmp2 <- frost$end_1st_dev_stage_days_since_may1st
days_since_may1st <- min(tmp1):max(tmp2)
rm(tmp1, tmp2)
par(mar = c(3, 3, .1, .1), mgp = c(2, .5, 0), tcl = -.3)
plot(frost$year, frost$bud_burst_days_since_may1st, type = "n",
     ylim = range(days_since_may1st), bty = "n")
for (index in 1:nrow(frost)) {
  tmp_x <- rep(frost$year[index], times = 2)
  tmp_y <- c(frost$bud_burst_days_since_may1st[index], 
             frost$end_1st_dev_stage_days_since_may1st[index])
  lines(x = tmp_x, y = tmp_y)
}
```

## Example of a `for` loop with `if`

Preparations:

```{r}
res <- data.frame(days_since_may1st = days_since_may1st, 
                  n_at_risk = NA)
```

### Illustrating the loop index:

```{r}
paint <- colorspace::divergingx_hcl(n = nrow(res), pal = "Zissou")
par(mar = c(3, 3, .1, .1), mgp = c(2, .5, 0), tcl = -.3)
plot(1:nrow(res), res$days_since_may1st, col = paint, pch = 16, bty = "n", 
     xlab = "for-loop index 'index'", ylab = "res$days_since_may1st at 'index'")
```

### An iteration 'by hand':

```{r}
par(mar = c(3, 3, .1, .1), mgp = c(2, .5, 0), tcl = -.3)
plot(frost$year, frost$bud_burst_days_since_may1st, type = "n",
     ylim = range(days_since_may1st), bty = "n")
for (index in 1:nrow(frost)) { ## here, the uninteresting loop
  tmp_x <- rep(frost$year[index], times = 2)
  tmp_y <- c(frost$bud_burst_days_since_may1st[index], 
             frost$end_1st_dev_stage_days_since_may1st[index])
  lines(x = tmp_x, y = tmp_y)
}
index <- 1
abline(h = res$days_since_may1st[index], lty = 2)
## boolean 1 and 2:
bool1 <- frost$bud_burst_days_since_may1st <= days_since_may1st[index]
bool2 <- frost$end_1st_dev_stage_days_since_may1st >= days_since_may1st[index]
which_true <- which(bool1 & bool2)
points(frost$year[which_true], days_since_may1st[index], col = paint[index], pch = 16)
res$n_at_risk[index] <- length(which_true)
rm(index)
```

### A 'full' loop:

```{r}
par(mar = c(3, 3, .1, .1), mgp = c(2, .5, 0), tcl = -.3)
plot(frost$year, frost$bud_burst_days_since_may1st, type = "n",
     ylim = range(days_since_may1st), bty = "n")
for (index in 1:nrow(frost)) { ## here, the uninteresting loop
  tmp_x <- rep(frost$year[index], times = 2)
  tmp_y <- c(frost$bud_burst_days_since_may1st[index], 
             frost$end_1st_dev_stage_days_since_may1st[index])
  lines(x = tmp_x, y = tmp_y)
}
for (index in 1:nrow(res)) { ## here, the interesting loop
  abline(h = res$days_since_may1st[index], lty = 2)
  bool1 <- frost$bud_burst_days_since_may1st <= days_since_may1st[index]
  bool2 <- frost$end_1st_dev_stage_days_since_may1st >= days_since_may1st[index]
  ## if any ... else ...
  if (any(bool1 & bool2)) {
    which_true <- which(bool1 & bool2)
    points(frost$year[which_true], 
           rep(days_since_may1st[index], times = length(which_true)), 
           col = paint[index], pch = 16)
    res$n_at_risk[index] <- length(which_true)
  } else {
    res$n_at_risk[index] <- 0
  }
}
head(res, n = 10)
```

```{r}
par(mar = c(3, 3, .1, .1), mgp = c(2, .5, 0), tcl = -.3)
plot(res$days_since_may1st, res$n_at_risk, type = "o", pch = 16, bty = "n")
grid()
```

<!--
```{r}
vektor <- levels(dd$d_cut)
par(mfrow = c(2, 5)) ## Ziel: immer zwei Kategorien zusammenpacken.
laenge <- 0 ## Hilfsobjekt: Starte mit 0.
for (index in vektor) {
  if (laenge == 0) {
    index_vektor <- index
    laenge <- 1 ## Hilfsobjekt: Setze auf 1.
  } else {
    index_vektor <- c(index_vektor, index)
    d_hier <- subset(d, d_cut %in% index_vektor)
    levels_kombiniert <- paste(index_vektor, collapse = " und ")
    plot(d_hier$d/100, d_hier$h, xlim = range(d$d)/100, ylim = range(d$h),
         pch = 16, col = rgb(0, 0.38, 0.27, alpha = 0.5), bty = "n",
         main = levels_kombiniert, xlab = "Stammdurchmesser [m]",
         ylab = "Baumh?he [m]", las = 1)
    cat(paste0("Plot f?r ", levels_kombiniert, " erstellt.\n"))
    laenge <- 0 ## Hilfsobjekt: Setze auf 0.
  }
}
```
-->

## 'while'-loops.

'while' loops are used less often in data management / preparation, but are more likely to be found in 'computationally intensive' applications (e.g. for optimizations).

**Nutzung:**
```{r, eval=F}
index <- k ## 'k' here has to be smaller than 'K' in next line.
while (index < K){
  ...
  index <- index + 1
}
```

- `...` und die darauf folgende Zeile wird so lange wiederholt wie die Bedingung `TRUE` ist (also solange hier `k`$<$`K`)
- flexible Anzahl an Wiederholungen.
- stoppt unmittelbar nachdem die Bedingung nicht mehr eingehalten wird.
- kann auch als eine 'for'-Schleife umgeschrieben werden.

- The commands that '`...` stands for, and the following line, are repeated as long as the condition is `TRUE` (i.e. here as long as `k`$<$`K`).
- flexible number of repetitions.
- stops immediately after the condition is no longer met.
<!--- can also be rewritten as a 'for' loop.-->

**Example 1**

```{r}
accepted <- 0
table(frost$n_frost > .5)
sum(frost$n_frost > .5)
P <- NULL
n_iter <- 0
while (accepted < 1000) {
  p <- rbeta(n = 1, shape1 = 1/3, shape2 = 1/3)
  y_tilde <- sample(x = c(TRUE, FALSE), size = nrow(frost), replace = T, 
                    prob = c(p, 1 - p))
  if (sum(y_tilde) == sum(frost$n_frost > .5)) {
    accepted <- accepted + 1
    P <- c(P, p)
  }
  n_iter <- n_iter + 1
}
length(P)
n_iter
length(P) / n_iter
summary(P)
```

**Example 2**

```{r}
set.seed(123)
x1 <- drought$elev - mean(drought$elev)
x2 <- runif(nrow(drought), min = min(x1), max = max(x1))
x2 <- x2 - mean(x2)
y <- drought$bair#  - mean(drought$bair)
f_y_work <- function(y, x1, x2, b0, b1, b2){-1 * (-2*y + 2*(b0 + b1*x1 + b2*x2))}
b0 <- 0
b1 <- 0
b2 <- 0
krit_diff <- 1 ## Initialisierung irgendwie so dass Bedingung am Anfang wahr ist.
krit_alt <- sqrt(mean(c(y - (b0 + b1*x1 + b2*x2))^2))
component <- NULL
while (krit_diff > 0.0001) { ## Beginn der while-Schleife.
  y_work <- f_y_work(y = y, x1 = x1, x2 = x2, 
                     b0 = b0[length(b0)], b1 = b1[length(b1)], 
                     b2 = b2[length(b2)])
  lm_b0 <- lm(y_work ~ 1)
  lm_b1 <- lm(y_work ~ -1 + x1)
  lm_b2 <- lm(y_work ~ -1 + x2)
  krit_b0 <- mean(lm_b0$residuals^2)
  krit_b1 <- mean(lm_b1$residuals^2)
  krit_b2 <- mean(lm_b2$residuals^2)
  selected <- which.min(c(krit_b0, krit_b1, krit_b2))
  update_weight <- rep(0, 3)
  update_weight[selected] <- .01
  b0 <- c(b0, b0[length(b0)] + update_weight[1] * coef(lm_b0))
  b1 <- c(b1, b1[length(b1)] + update_weight[2] * coef(lm_b1))
  b2 <- c(b2, b2[length(b2)] + update_weight[3] * coef(lm_b2))
  component <- c(component, selected)
  krit_neu <- sqrt(mean(c(y - (b0[length(b0)] + 
                                 b1[length(b1)] * x1 + 
                                 b2[length(b2)] * x2))^2))
  krit_diff <- krit_alt - krit_neu ## Update!
  krit_alt <- krit_neu
} ## Ende der while-Schleife.
table(component)
par(mfrow = c(3, 1), mar = c(3, 3, 0, 0), las = 1, oma = c(0, 0, 0, 0), 
    mgp = c(2, .4, 0), tcl = -.3)
paint <- colorspace::divergingx_hcl(n = length(b0), pal = "Zissou")
paint_a <- colorspace::divergingx_hcl(n = length(b0), pal = "Zissou", alpha = .1)
plot(x1, y, pch = 16, bty = "n", las = 1, ylim = range(c(0, y)), bty = "n")
for (index in 1:length(b0)) {
  abline(a = b0[index], b = b1[index], col = paint_a[index])
}
plot(x2, y, pch = 16, bty = "n", las = 1, ylim = range(c(0, y)), bty = "n")
for (index in 1:length(b0)) {
  abline(a = b0[index], b = b2[index], col = paint_a[index])  
}
plot(as.numeric(as.factor(component)), yaxt = "n", ylab = "Component", 
     col = paint, pch = 16, bty = "n")
axis(2, at = 1:length(unique(component)), labels = levels(as.factor(component)), 
     las = 1)
```

<!--
% 'repeat'-Schleifen.}
% Nutzung:
% \begin{Verbatim}
% repeat{
%   CONTENT
%   if(CONDITION)
%     break
% }
% \end{Verbatim}
% 
% - theoretically endless repetition of a task
% - parameter changes need to be done in the loop
% - termining condition needed
% - `break} ends the current subroutine
-->

## 'apply'-commands

An 'apply'-command applies the same function to each of the elements of a data object.

**Usage:**
```{r, eval=F}
apply(X, MARGIN, FUN, ...) ## For matrix X: Result is a list.
lapply(X, FUN, ...) ## For list X: Result is a list.
sapply(X, FUN, ...) ## For list X: Result is a vector or another 
                    ## Data object that the result might be 'simplified' to.
```

- `apply` applies function (specified by `FUNCTION`) to each element of the respective dimension (defined with argument `MARGIN`) of `X`.
- `MARGIN` equals `1` for line-by-line, and `2` for column-wise execution.
- `...` for further arguments to `FUNCTION` (same for every element of` X`!).
- For lists `X`, `MARGIN` cannot be selected because lists only have one dimension.

### Exercises

```{r, warning=FALSE}
apply(drought, MARGIN = 2, FUN = function(x){sum(is.na(x))})
apply(drought[, 1:2], MARGIN = 2, FUN = mean)
apply(drought[, 1:2], MARGIN = 1, FUN = mean)
apply(frost, MARGIN = 2, FUN = function(x){sum(is.na(x))})
apply(frost[, c(1:2, 6:7)], MARGIN = 2, FUN = mean)
lapply(frost[, c(1:2, 6:7)], FUN = mean)
sapply(frost[, c(1:2, 6:7)], FUN = mean)
```

## Programming-'Workflow'.

- Use loops as often as possible ('upwards!'), but avoid loops as often as necessary ('downwards'), because (very roughly (!) said):
   - Loops read and write to the main memory in each iteration $\rightarrow$.
   - Vectorized programming reads and writes only once: many functions take vectors as arguments and are therefore (often) faster.
- For clearer code:
   - vectorizing conditions:
```{r, eval=F}
result <- content1 * logical_comparison1 +
          content2 * logical_comparison2 + ...
```

- Use an `apply` command if you want the function to do the same on every element.
- But:
   - Loops are easy and whoever masters them is already a king: It is better if R-Code gets something right slowly than quickly wrong!
   - Loops cannot be avoided in iterative processes!
   - Avoiding 'if-else' is not worth it under complex conditions!

`r if (knitr::is_latex_output()) '\\clearpage'`

# Define your own functions.

**Why should I be able to define my own functions?**

- Functions generalize command sequences and make it easier and easier to try something out under many different argument values / dates / ....
- Functions keep the workspace clean (see next section on environments).
- Functions facilitate the reproducibility of analyzes.
- Functions make it easier for other users to access your work.
- As can be seen from the `apply()` examples, it is very often necessary to be able to write your own little helper functions.
Also for your own **orientation**: Always comment on the processes and steps in your code and in your functions to make it easier to understand the motivation and ideas behind it later.

```{r, eval=F}
name <- function(arg1, arg2, arg3 = TRUE, arg4 = 2, ...){
  content
  return(result)
}
```

- The general rules for naming objects also apply to function arguments.
- Arguments can have preset values (here `arg3` and` arg4`)
- The last argument `...` (optional) is a special argument and can be used to pass unspecified arguments to function calls.
- Arguments changed by `content` and objects created are in their own local environment.
- The result is returned to the global environment with `return(result)`.

## Naming conventions for arguments.

Argument name  Inhalt
-------------- -------
`data`         Dataframe
`x`, `y`, `z`  Vectors (most often with numerical elements)
`n`            Sample size
`formula`      Formula object
...            ...

- Use function and argument names that are based on existing R functions.
- Make arguments as self-explanatory as possible by name.

<!--
% functions: options for function attributes}
% - functions without attributes are possible (eg. `getwd()}, `ls()})
% - attributes without default values have to be filled in
% - default values are used if attributes is not specified in the call to the function
-->

## `content` and` result`.

**The `content` block:**

- Should make it possible to carry out many similar -- but different -- calculations and therefore define as few objects as possible to 'fixed values': alternatively, always try to define arguments with default values.
- Falls back on the higher-level environment (or environments, if necessary) if it cannot find an object in the local environment (this is known as *scoping*).

**The `result` object:**

- Can be of any possible R object class (vector, list, data set, function (a function that itself returns a function is called *closure*), ...).
- Is generated by calling the function and stored in the global environment.
- All other objects are no longer 'visible' from the global environment.

## Examples

### Environments and Scoping

```{r, error=TRUE}
rm(list = ls())
ls()
f <- function(x){
  y <- 2
  print(ls())
  y <- y + z ## f wird nach z in der ?bergeordneten Umgebung (hier global) suchen.
  print(ls())
  return(x + y)
}
x <- 1; z <- 3
f(x = x) ## f wird z finden:
## -> obwohl wir es nicht explizit als Argument in die lokale Umgebung
##    von f übergeben haben.
y ## Von der übergeordneten Umgebung aus können wir nicht auf y zurückgreifen.
```

### Closure

```{r}
power <- function(exponent){
  return(function(x){
    return(x ^ exponent)})
}
square <- power(2)
square(2)
square(4)
cube <- power(3)
cube(2)
cube(4)
```

<!--
%functions: recursions}
%\begin{Verbatim}
%fibonacci <- function(n){
%  if(n == 1)
%    return(0)
%  else if(n == 2)
%    return(1)
%  else
%    return(fibonacci(n-1) + fibonacci(n-2))
%}
%\end{Verbatim}
%
%- call the function from within, define break point
%- easy for small tasks, memory intensive for large data
% functions: numerical minimization}
%\begin{Verbatim}
%goal <- function(par,...){
%  CONTENT
%  return(value)
%}
%nlm(par,goal,...)
%\end{Verbatim}
%
%- find parameter of function that produces lowest value of goal function
%- alternative: `optim}
%- if parameter constrained to interval, use: `optimize}
%- uses grid search to find minimum
%- use to find: optimal smoothing parameter, maximum likelihood, etc.

%% 
%functions: finding and handling errors}
%
%\begin{tabular}{r|l}
%Code & Function \\ \hline
%`traceback} & shows subroutine and line where error occurred \\
%`browser} & halts function and opens interactive console \\
%`debug} & execute function completely in browser mode \\
%`try} & wraps a part of code that is likely to produce errors \\
%\end{tabular}
%
%- missing brackets or syntax errors are reported while reading the function
%- the more complex the function the less obvious is any fail
%- if an analysis takes longer, multiple attempts on solving a problem are not efficient anymore
%- otherwise insert 'print' of object that might contain error
%
-->

`r if (knitr::is_latex_output()) '\\clearpage'`

`r if (knitr::is_html_output()) '# References {-}'`
